Bird beak's elaboration and exploration

### 2021-10-19 Analysis/03-elaboration_exploration_analyses.Rmd
Update 03-elaboration_exploration_analyses.Rmd

```{r, eval = TRUE, echo = FALSE, print = FALSE}
library(beer)
## Loading the MCMCglmm
load("../Data/processed/combined_results.rda")
## Loading the tree and data
data(demo_data)
data(demo_tree)
## Setting the colour vector
colour_vector <- c("orange", "blue", "darkgreen", "grey")
```

@@

```{r, eval = TRUE, echo = FALSE}
covar_matrices <- MCMCglmm.dispRity(
                  data          = demo_data,
                  posteriors    = combined_results,
                  group         = MCMCglmm.levels(combined_results)[1:4],
                  rename.groups = c("gulls", "plovers", "sandpipers", "phylogeny"))
```

```{r, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 5, echo = FALSE}
## Plotting just the space
sauron.plot(covar_matrices,
            col = colour_vector,
            main = "demo space",
            legend = TRUE)
```

@@

```{r, echo = FALSE, fig.height = 6, fig.width = 6}
sauron.plot(covar_matrices,
            col = colour_vector,
            main = "Average ellipses and\nmajor axes",
            major.axes = mean,
            ellipses = mean,
            legend = TRUE,
            cex = 0.5)
```

@@

```{r, echo = FALSE, fig.height = 12, fig.width = 6}
par(mfrow = c(2,1))
## Plotting 100 axes
sauron.plot(covar_matrices,
            n = 100,
            col = colour_vector,
            main = "100 random major axes",
            major.axes = TRUE,
            legend = TRUE,
            points = FALSE,
            xlim = c(-2, 2),
            ylim = c(-2, 2))
## Plotting 100 ellipses
sauron.plot(covar_matrices,
            n = 100,
            col = colour_vector,
            main = "100 random ellipses",
            ellipses = TRUE,
            legend = TRUE,
            points = FALSE,
            xlim = c(-2, 2),
            ylim = c(-2, 2))
```

@@

```{r, fig.height = 6, fig.width = 18, echo = FALSE}
## Run the analyses between groups
results_between <- covar.projections.wrapper(covar_matrices,
                                             type = "between",
                                             base = "phylogeny",
                                             dispRity.out = TRUE)
## Plot the results
wrap.plot <- function(results, col, main) {
    ## How many sub plots?
    n <- length(results)
    ## Which will be the centre plot?
    centre <- ceiling(n/2)
    ## Plot everything
    par(mfrow = c(1, n))
    for(i in 1:n) {
        plot(results[[i]], col = col, main = ifelse(i == centre, main, ""))
    }
}
wrap.plot(results_between,
          col = colour_vector[-4],
          main = "Groups difference")
```
Figure: these results show the projections ("position"), rejection ("distance") and angles ("degree") of the posteriors of each group's major axes onto the corresponding global phylogenetic major axes. For each panel, the highest the score the more the clade is elaborating, exploring, exploring perpendicular to the main axes for the position, distance and degree respectively.

@@

```{r, fig.height = 6, fig.width = 18, echo = FALSE, eval = TRUE}
## Run the analyses of groups on the phylogeny
results_phylo <- covar.projections.wrapper(covar_matrices,
                                            type = "within",
                                            base = "phylogeny",
                                            dispRity.out = TRUE)
## Plot the results
wrap.plot(results_phylo,
          col = colour_vector[-4],
          main = "Projection on the\nglobal phylogeny axes")
```
Figure: the projection of each species per group onto the posterior distribution of global major phylogenetic axis.

@@

```{r, fig.height = 6, fig.width = 18, echo = FALSE, eval = TRUE}
## Run the analyses groups on themselves
results_within <- covar.projections.wrapper(covar_matrices,
                                            type = "within",
                                            dispRity.out = TRUE)
## Plot the results
wrap.plot(results_within,
          col = colour_vector,
          main = "Projection on the\ngroup's phylogeny axes")
```
Figure: the projection of each species per group onto the posterior distribution of each clade's major phylogenetic axis (the grey distributions corresponds to the projections of all the species, irregardless of their group, onto the global phylogenetic major axes).

```{r, echo = FALSE, eval = TRUE}
library(phytools)
source("~/Packaging/phytools/R/plotTree.wBars.R")

get.disparity.matrix <- function(data, col) {
    ## Get the matrices
    matrices <- lapply(data$disparity, `[[`, 1)
    ## Add the rownames
    list_of_names <- list()
    for(one_group in 1:length(matrices)) {
        rownames(matrices[[one_group]]) <- list_of_names[[one_group]] <- rownames(data$matrix[[1]])[c(data$subsets[[one_group]]$elements)]
    }
    ## Name the groups
    names(list_of_names) <- names(data$subsets)

    ## Matrix
    matrix <- do.call(rbind, matrices)

    ## Colours
    if(!missing(col)) {
        name.cols <- function(names, col) {
            out <- rep(col, length(names))
            names(out) <- names
            return(out)
        }
        col <- mapply(name.cols, list_of_names, as.list(col))
        names(col) <- NULL
        col <- unlist(col)
    }

    ## Output
    if(missing(col)) {
        return(list(data = matrix, names = list_of_names))
    } else {
        return(list(data = matrix, names = list_of_names, col = col))
    }
}

## Get the different elements
phylo_elaboration  <- get.disparity.matrix(results_phylo$position,
                                          col = colour_vector[-4])
within_elaboration <- get.disparity.matrix(
    get.subsets(results_within$position, subsets = c(1,2,3)),
                                          col = colour_vector[-4])
phylo_exploration  <- get.disparity.matrix(results_phylo$distance,
                                          col = colour_vector[-4])
within_exploration <- get.disparity.matrix(
    get.subsets(results_within$distance, subsets = c(1,2,3)),
                                          col = colour_vector[-4])
phylo_angle        <- get.disparity.matrix(results_phylo$degree,
                                          col = colour_vector[-4])
within_angle       <- get.disparity.matrix(
    get.subsets(results_within$degree, subsets = c(1,2,3)),
                                          col = colour_vector[-4])
```

>>>>

```{r, eval = TRUE, echo = FALSE, print = FALSE}
library(dispRity)
## Loading the tree and data
data(charadriiformes)
## Setting the colour vector
colour_vector <- c("orange", "blue", "darkgreen", "grey")
```

@@

```{r, eval = TRUE, echo = FALSE}
covar_matrices <- MCMCglmm.subsets(
                  data          = charadriiformes$data,
                  posteriors    = charadriiformes$posteriors,
                  group         = MCMCglmm.levels(charadriiformes$posteriors)[1:4],
                  rename.groups = c("gulls", "plovers", "sandpipers", "phylogeny"))
```

```{r, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 5, echo = FALSE}
## Plotting just the space
covar.plot(covar_matrices,
            col = colour_vector,
            main = "demo space",
            legend = TRUE)
```

@@

```{r, echo = FALSE, fig.height = 6, fig.width = 6}
covar.plot(covar_matrices,
            col = colour_vector,
            main = "Average ellipses and\nmajor axes",
            major.axes = mean,
            ellipses = mean,
            legend = TRUE,
            cex = 0.5)
```

@@

```{r, echo = FALSE, fig.height = 12, fig.width = 6}
par(mfrow = c(2,1))
## Plotting 100 axes
covar.plot(covar_matrices,
            n = 100,
            col = colour_vector,
            main = "100 random major axes",
            major.axes = TRUE,
            legend = TRUE,
            points = FALSE,
            xlim = c(-2, 2),
            ylim = c(-2, 2))
## Plotting 100 ellipses
covar.plot(covar_matrices,
            n = 100,
            col = colour_vector,
            main = "100 random ellipses",
            ellipses = TRUE,
            legend = TRUE,
            points = FALSE,
            xlim = c(-2, 2),
            ylim = c(-2, 2))
```

@@

```{r, fig.height = 6, fig.width = 18, echo = FALSE}
## Run the analyses between groups
results_between <- dispRity.covar.projections(covar_matrices,
                                             type = "groups",
                                             base = "phylogeny")
## Plot the results
wrap.plot <- function(results, col, main) {
    ## How many sub plots?
    n <- length(results)
    ## Which will be the centre plot?
    centre <- ceiling(n/2)
    ## Plot everything
    par(mfrow = c(1, n))
    for(i in 1:n) {
        plot(results[[i]], col = col, main = ifelse(i == centre, main, ""))
    }
}
wrap.plot(results_between,
          col = colour_vector[-4],
          main = "Groups difference")
```
Figure: these results show the projections ("position"), rejection ("distance") and angles ("degree") of the posteriors of each group's major axes onto the corresponding global phylogenetic major axes. For each panel, the highest the score the more the clade is elaborating, exploring, exploring perpendicular to the main axes for the position, distance and degree respectively.

@@

```{r, fig.height = 6, fig.width = 18, echo = FALSE, eval = TRUE}
## Run the analyses of groups on the phylogeny
results_phylo <- dispRity.covar.projections(covar_matrices,
                                            type = "elements",
                                            base = "phylogeny")
## Plot the results
wrap.plot(results_phylo,
          col = colour_vector[-4],
          main = "Projection on the\nglobal phylogeny axes")
# source("../Functions/circular.plot.R")

# circular.plot.phytools(data = results_phylo$position, tree = demo_tree, gradient = c("blue", "orange"), clade.names = TRUE, clades = colour_vector[-4], lwd = 3)
```
Figure: the projection of each species per group onto the posterior distribution of global major phylogenetic axis.

@@

```{r, fig.height = 6, fig.width = 18, echo = FALSE, eval = TRUE}
## Run the analyses groups on themselves
results_within <- dispRity.covar.projections(covar_matrices,
                                            type = "elements")
## Plot the results
wrap.plot(results_within,
          col = colour_vector,
          main = "Projection on the\ngroup's phylogeny axes")
```
Figure: the projection of each species per group onto the posterior distribution of each clade's major phylogenetic axis (the grey distributions corresponds to the projections of all the species, irregardless of their group, onto the global phylogenetic major axes).
